Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-09-16 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:09:16	22:56:55	(Wed16.Sep)

\author{Jo~Inge~Buskenes, %
        Roy~Edgar~Hansen,%
        Andreas~Austeng%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem All authors are with the Department
of Informatics, University of Oslo, Norway.}% <-this % stops a space

@@

In many cases MVDR can improve image contrast and resolution compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, the computational complexity of conventional beamformers are linear with the number of channels, O($M$), while MVDR is at O($M^3$). This makes MVDR impracticable and sometimes infeasible. We have shown in previous studies that MVDR's complexity can be reduced, as well as accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014,Asen2013}, but there is a way to avoid it altogether while achieving similar performance.

@@

In this work we aim to provide some intuitive understanding of how the window database may be designed. Like Synnev\aa{}g et. al. we limit our study to the use of variations of the parametric Kaiser window, since we observed negligible performance difference to using this function or any of the common trigonometric ones. However, we steer all our windows and not just some, and provide reason why the Kaiser window is a suitable one, what variations of Kaiser we should include, and how much it should be steered.

@@

\begin{figure*}[t]\centering
\subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_beta.pdf}%%
\label{oversampling_mosaic_bounds_beta}}
\hfil
\subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA prefers selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_phi.pdf}%
\label{oversampling_mosaic_bounds_phi}}
\caption{Kaiser windows chosen for each pixel in the image. The underlying image is the one shown in \Fig{oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
\label{fig_sim}
\end{figure*}

@@

What is the best set of Kaiser windows to use with LCA, and how much should they be steered? To investigate this we constructed large and equally sized window databases containing windows with varied upper boundaries for $\beta$ and $\phi$. As lower boundaries we chose $\beta=0$ and $\phi=0^\circ$, which includes the symmetric rectangularly weighted DAS image. The images produced are shown in \Fig{oversampling_mosaic_bounds}. In each image we computed the lateral distance $\Delta x$ between the -3dB points of the anchor, with the reference $\Delta x_\text{ref}$ being the rectangularly weighted DAS image. A forth order polynomial fit was used in this computation. This measure is also called the full width half maximum (FWHM). In the case of a imaging a point scatter, it is closely related to the resolution of of the system. For a rectangular window, for instance, the system resolution is $\Delta = \text{FWHM}/0.89$~\cite{Harris1978}.

>>>>

\author{Jo~Inge~Buskenes, %
        Roy~Edgar~Hansen, %
        Andreas~Austeng%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem All authors are with the Department
of Informatics, University of Oslo, Norway.}% <-this % stops a space

@@

In many cases MVDR can improve image contrast and resolution compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, the computational complexity of conventional beamformers are linear with the number of channels, O($M$), while MVDR is at O($M^3$). This makes MVDR impracticable and sometimes infeasible. We have shown in previous studies that MVDR's complexity can be reduced, as well as accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014,Asen2013}, but there is a way to avoid using MVDR altogether while achieving similar performance.

@@

In this work apply the LCA method to active sonar imaging, and aim to provide some intuitive understanding of how the window database may be designed. Like Synnev\aa{}g et. al. we limit our study to the use of variations of the parametric Kaiser window, since we observed negligible performance difference to using this function or any of the common trigonometric ones. However, we steer all our windows and not just some, and provide reason why the Kaiser window is a suitable one, what variations of Kaiser we should include, and how much it should be steered.

@@

\begin{figure*}[t]\centering
\subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_beta.pdf}%%
\label{oversampling_mosaic_bounds_beta}}
\hfil
\subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_phi.pdf}%
\label{oversampling_mosaic_bounds_phi}}
\caption{Kaiser windows chosen for each pixel in the image. The underlying image is the one shown in \Fig{oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
\label{fig_sim}
\end{figure*}

\begin{figure*}[t]\centering
\subfloat[]{\includegraphics[width=\textwidth]{gfx/beamformer_comparison_holmengraa.pdf}%%
\label{oversampling_mosaic_bounds_beta}}
\hfil
\subfloat[]{\includegraphics[width=\textwidth]{gfx/beamformer_comparison_cross.pdf}%
\label{oversampling_mosaic_bounds_phi}}
\caption{}\label{fig_sim}
\end{figure*}

@@

What is the best set of Kaiser windows to use with LCA, and how much should they be steered? To investigate this we constructed large and equally sized window databases containing windows with varied upper boundaries for $\beta$ and $\phi$. As lower boundaries we chose $\beta=0$ and $\phi=0^\circ$, which includes the symmetric rectangularly weighted DAS image. The images produced are shown in \Fig{oversampling_mosaic_bounds}. In each image we computed the lateral distance $\Delta x$ between the -3dB points of the anchor, with the reference $\Delta x_\text{ref}$ being the rectangularly weighted DAS image. A forth order polynomial fit was used in this computation. This measure is also commonly called the full width half maximum (FWHM). In the case of a imaging a point scatter, it is closely related to the resolution of of the system. For a rectangular window, for instance, the system resolution is approximately $\Delta \approx \text{FWHM}/0.89$~\cite{Harris1978}.
